### 2022-08-19 publications/2019ScalingOnTheEdge/img/legoflow.tex

\adjustbox{scale=0.65}{
    \ttfamily\centering%\fbox{%
    \begin{tikzpicture}[align=center,
    node distance=0.7cm and 0.7cm,
    every initial by arrow/.style={draw=none, minimum size=0em}]
    % Place nodes       
    \node [state, draw=none, minimum size=0em] (input) {};       
    \node [state, right=1cmof input, minimum size=7em] (detect) {Detection};               
    \node [state, right=2cm of detect, minimum size=7em] (symb) {Symbolic\\Repr.};
    \node [state, right=.5cm of symb, minimum size=7em] (feedback) {Feedback\\Gen.};
    \node [state, accepting, draw=none, right=1cm of feedback, minimum size=0em] (output1) {};
    \node [state, accepting, draw=none, below=1.4cm of output1, minimum size=0em] (output2) {};

    % Draw edges
    \path[draw, -{Latex[length=3mm]}]
    (input) edge node[above left] {Input} (detect)
    (detect) edge node[above] {LEGO bricks\\in input} (symb)
    (symb) edge node {} (feedback)
    (feedback) edge node[above right] {Feedback} (output1)
    (detect) edge [bend right=10] node[below left] {No LEGO bricks, thus no feedback} (output2) ;
    \end{tikzpicture}
}%}