### 2022-08-19 publications/2019ScalingOnTheEdge/img/pipeline.tex

\adjustbox{scale=0.65}{
        \ttfamily\centering%\fbox{%
        \begin{tikzpicture}[align=center,
                        node distance=0.7cm and 0.7cm,
                        every initial by arrow/.style={draw=none, minimum size=0em},
                        %node styles:
                        block_center/.style ={rectangle, draw=black, thick, fill=white,
                                        text width=8em, text centered, minimum height=4em},
                        block_left/.style ={rectangle, draw=black, thick, fill=white,
                                        text width=16em, text ragged, minimum height=4em, inner sep=6pt},
                        block_noborder/.style ={rectangle, draw=none, thick, fill=none,
                                        text width=18em, text centered, minimum height=1em},
                        block_assign/.style ={rectangle, draw=black, thick, fill=white,
                                        text width=18em, text ragged, minimum height=3em, inner sep=6pt},
                        block_lost/.style ={rectangle, draw=black, thick, fill=white,
                                        text width=16em, text ragged, minimum height=3em, inner sep=6pt},
                        block_rounded/.style ={rectangle, draw=black, thick, fill=white,
                                        text width=8em, text centered, rounded corners=.55cm, minimum height=4em},
                        line/.style ={draw, thick, -{Latex[length=3mm]}, shorten >=0pt}
                ]

                \matrix [column sep=5mm,row sep=3mm] (top) {
                        \node [block_center] (detect) {Detection};
                         & \node [block_center] (repr) {Symbolic\\Repr.};
                         & \node [block_center] (model) {Task Model\\$M\{S, E\}$};
                         & \node [block_center] (feedback) {Feedback\\Generation}; \\[3ex]
                };

                \matrix [column sep=13mm,row sep=3mm, below=of top] (bottom) {
                        \node [block_rounded] (sensors) {On-body\\Sensors};
                         %& \node [block_rounded] (user) {Human User};
                         %&\node[cloud, draw, black, text centered, cloud ignores aspect=false] (user) {Human User};
                         & \node[maninblack, minimum height=5em] (user) {\Large Human User};
                         & \node [block_rounded] (hud) {HUD, Speakers,\\etc.}; \\
                };

                %\Smiley[1, below of=bottom] (smiley);

                \begin{scope}[every path/.style=line]
                        \path (detect) -- (repr);
                        \path (repr) -- (model);
                        \path (model) -- (feedback);
                        \path (feedback) -- (hud);
                        \path (hud) -- (user);
                        \path (user) -- (sensors);
                        \path (sensors) -- (detect);
                \end{scope}

                % % Place nodes       
                % \node [initial, state, draw=none, initial text=, minimum size=0em] (input) {};
                % \node [state, right=1cmof input, minimum size=7em] (detect) {Detection};
                % \node [state, right=2cm of detect, minimum size=7em] (symb) {Symbolic\\Repr.};
                % \node [state, right=.5cm of symb, minimum size=7em] (feedback) {Feedback\\Gen.};
                % \node [state, accepting, draw=none, right=1cm of feedback, minimum size=0em] (output1) {};
                % \node [state, accepting, draw=none, below=1.4cm of output1, minimum size=0em] (output2) {};

                % % Draw edges
                % \path[draw, -{Latex[length=3mm]}]
                % (input) edge node[above left] {Input} (detect)
                % (detect) edge node[above] {LEGO bricks\\in input} (symb)
                % (symb) edge node {} (feedback)
                % (feedback) edge node[above right] {Feedback} (output1)
                % (detect) edge [bend right=10] node[below left] {No LEGO bricks, thus no feedback} (output2) ;
        \end{tikzpicture}
}%}