### 2022-08-19 publications/2019EdgeDroid/img/taskmodel.tex

\adjustbox{scale=0.75}{
    \ttfamily\centering%\fbox{%
    \begin{tikzpicture}[align=center,
        node distance=.5cm and 1cm,
        every initial by arrow/.style={-{Latex[length=2mm]}},
        line/.style ={draw, thick, -latex', shorten >=0pt}]
        % Place nodes       
        \node [state, minimum size=3em] (si) {\Large $s_{i}$};
        \node [state, above right=of si, minimum size=3em] (sj) {\Large $s_{j}$};
        \node [state, below right=of si, minimum size=3em] (se) {\Large $s^{e}_{k}$};

        \path[draw, -{Latex[length=2mm]}]
        (si)
        edge [out=140,in=220,looseness=6] node[left] {\emph{In:} none\\\emph{Out:} positive feedback} (si)
        edge [bend left=20] node[above left] {\emph{In:} correct action\\\emph{Out:} none} (sj)
        edge [bend left=20] node[right] {\emph{In:} incorrect action\\\emph{Out:} none} (se)

        (se)
        edge [bend left=20] node[below left] {\emph{In:} undo action\\\emph{Out:} none} (si)
        edge [out=230,in=310,looseness=6] node[below] {\emph{In:} none\\\emph{Out:} negative feedback} (se)

        (sj) edge [out=50,in=130,looseness=6] node[above] {\emph{In:} none\\\emph{Out:} positive feedback} (sj)
        ;

    \end{tikzpicture}
}%}