### 2022-08-19 publications/2021ImpactDelayedResponse/plots/lego_hierarchy_diagram.tex

\begin{tikzpicture}[grow=down, scale=.6]
    \tikzset{edge from parent/.style= 
            {draw, thick, edge from parent fork down},
            every tree node/.style=
            {draw,very thick,align=center,scale=1,fill=black!10!white},
            every leaf node/.style=
            {fill=green!60!black!25!white,text width=2cm,minimum width=2cm},
            level distance=1in,sibling distance=.05in,
            execute at begin node=\strut
            }

    \node[draw,dashed,align=center](command){
        \footnotesize{Example instruction:}\\
        \small\textbf{``Find a $1 \times 1$ black piece and add it to}\\
        \textbf{the top right of the current model.''}
    % \\
    % \large{\textbf{``Remove the $1 \times 3$ red piece from the top of the current model.''}}
    };

    \begin{scope}[shift={($(command.south)+(0,-.25in)$)}]
        \Tree [.\node[fill=red!50!white]{\textbf{\Large{Change model}}};
                \node[diamond, aspect=2, fill=yellow!25!white](choice){Add or remove a piece?};
                ]
    \end{scope}

    \begin{scope}[shift={($(choice.west)-(1.5in,0)$)}]
        \Tree [.\node[draw=none, fill=none](add){Add a piece};
                [.{Select piece $X$} 
                    {Set $X$ size and color}
                    {Search box for matching pieces} ]
                [.{Select location $Y$} 
                    {Set $Y$ vertical and horizontal position}
                    {Find location on model} ]
                [.{Place $X$ at $Y$} 
                    {Pick up $X$ and move to model}
                    {Attach $X$ to model} ]
            ]
    \end{scope}

    \begin{scope}[shift={($(choice.east)+(1.5in,0)$)}]
        \Tree [.\node[draw=none, fill=none](remove){Remove a piece};
                \node[draw=none, fill=none]{\Huge{\ldots}}; 
                % [.{Select location $Y$} 
                %     {Set $Y$\\vertical and\\horizontal\\position}
                %     {Find\\location\\on model} ]
                % [.{Select piece $X$ at $Y$}
                %     {Set $X$\\size and\\color}
                %     {Search $Y$ for\\matching\\pieces} ]
                % [.{Remove piece $X$} 
                %     {Detach\\$X$}
                %     {Return $X$\\to box} ]
            ]
    \end{scope}

    \draw [thick] (choice.west) -- (add.east);
    \draw [thick] (choice.east) -- (remove.west);

\end{tikzpicture}%