### 2022-08-31 publications/2022Ainur/abstract.tex
cleave and ainur

\begin{abstract}
  Experimental research on wireless networking in combination with edge and cloud computing has been the subject of explosive interest in the last decade.
  This development has been driven by the increasing complexity of modern wireless technologies and the  extensive softwarization of these through projects such as a \acf*{O-RAN}.
  In this context, a number of small- to mid-scale testbeds have emerged, employing a variety of technologies to target a wide array of use-cases and scenarios in the context of novel mobile communication technologies such as 5G and beyond-5G.
  Little work, however, has yet been devoted to developing a standard framework for wireless testbed automation which is hardware-agnostic and compatible with edge- and cloud-native technologies.
  Such a solution would simplify the development of new testbeds by completely or partially removing the requirement for custom management and orchestration software.

  In this paper, we present the first such mostly hardware-agnostic wireless testbed automation framework, \emph{Ainur}.
  It is designed to configure, manage, orchestrate, and deploy workloads from an end-to-end perspective.
  Ainur is built on top of cloud-native technologies such as Docker, and is provided as \acs*{FOSS} to the community through the {KTH-EXPECA/Ainur} repository on GitHub.
  We demonstrate the utility of the platform with a series of scenarios, showcasing in particular its flexibility with respect to physical link definition, computation placement, and automation of arbitrarily complex experimental scenarios.
\end{abstract}