### 2022-08-31 publications/2022Ainur/paper.tex

\documentclass[conference,10pt,letterpaper,twocolumn]{IEEEtran}

% begin packages
% ==============
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[htt]{hyphenat}  % for linebreaks in texttt
\usepackage[pdftex]{graphicx}
% \graphicspath{{./images/}{./plots/}{./figures}}
% \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\usepackage{amsmath}
\usepackage{url}
\usepackage{orcidlink}
\usepackage[inline]{enumitem}
\setlist[description]{font=\itshape\underline}
\usepackage{todonotes}
\usepackage{caption}
\usepackage{subcaption}
% \usepackage[newfloat=true]{minted}
% \setminted{breaklines=true, frame=lines}
% \usemintedstyle{xcode}
% \SetupFloatingEnvironment{listing}{name=List.,within=none}
% \captionsetup[table]{justification=centerlast,
%                      labelsep=newline,
%                      font=sf,
%                      textfont=footnotesize}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage[detect-weight=true,detect-family=true,binary-units=true,list-units=single,range-units=single]{siunitx}
\usepackage[nolist]{acronym}
\input{acronyms.tex}

% references
\usepackage[
  style=numeric-comp,
  sorting=none,
  sortcites,
  hyperref,
  mincitenames=1,
  maxcitenames=2,
  maxbibnames=2,
  minbibnames=1,
  citestyle=numeric-comp, % for [1, 2] instead of [1], [2]
  backend=bibtex
]{biblatex}
\bibliography{references.bib}
\AtBeginBibliography{\small}
\AtEveryBibitem{\clearfield{day}}
\AtEveryBibitem{\clearfield{isbn}}
% \AtEveryBibitem{\clearfield{url}}
\AtEveryBibitem{\clearfield{series}}
\AtEveryBibitem{\clearlist{location}}
\AtEveryBibitem{\clearfield{doi}}

% for dealing with ACM's stupid bibtex requirements
% \usepackage{biblatex2bibitem}
% \usepackage[nosort,compress]{cite}

\usepackage{easyReview}
\usepackage{outline}
\usepackage{orcidlink}
\usepackage[all]{hypcap}
\usepackage{makecell}
\usepackage[capitalize,nameinlink,noabbrev]{cleveref}
% \crefname{subfigure}{Subfigure}{Subfigures}
% \Crefname{subfigure}{Subfigure}{Subfigures}
% \crefname{listing}{Listing}{Listings}
% \Crefname{listing}{Listing}{Listings}

\hypersetup{
  hidelinks,
  colorlinks=true,
  allcolors=black,
  pdfstartview=Fit,
  breaklinks=true
}

\newcommand{\kthemail}[1]{\href{mailto:#1@kth.se}{#1}}

\begin{document}
\title{Ainur:~A Framework for Repeatable End-to-End Wireless Edge Computing Testbed Research}

\author{%
  \IEEEauthorblockN{%
  Manuel {Olguín Muñoz}\IEEEauthorrefmark{1}~\orcidlink{0000-0002-3383-2335}, Seyed Samie Mostafavi\IEEEauthorrefmark{2}~\orcidlink{0000-0001-9316-0414}, Vishnu N. Moothedath\IEEEauthorrefmark{3}~\orcidlink{0000-0002-2739-5060}, James Gross\IEEEauthorrefmark{4}~\orcidlink{0000-0001-6682-6559}%
  }%
  \IEEEauthorblockA{%
    School of Electrical Engineering \& Computer Science\\%
    KTH Royal Institute of Technology, Sweden\\%
    Email:~\{\IEEEauthorrefmark{1}\kthemail{molguin}, \IEEEauthorrefmark{2}\kthemail{ssmos}, \IEEEauthorrefmark{3}\kthemail{vnmo}, \IEEEauthorrefmark{4}\kthemail{jamesgr}\}\href{mailto:molguin@kth.se}{@kth.se}
    % \IEEEauthorrefmark{1}Corresponding author.%
  }
}
\maketitle

\input{abstract.tex}
\input{introduction}
\input{ainur}
\input{demo}
\input{conclusion}
\input{acks.tex}
% \input{futurework}

\printbibliography{}
\end{document}