### 2022-08-31 publications/2022Ainur/software_layers/workload.tex

\subsubsection{Workloads}\label{sec:workload}

\begin{listing}[tb]
\caption{Workload specification example.}\label{lst:compose}
\begin{minted}{yaml}
name: FooBar
duration: '3h 15m'
version: '1.2a'
# ...
compose:
  version: '3.9'
  services:
  fooServer:
    image: 'foo:server'
    hostname: 'server.{{.Task.Slot}}'
    command: '/usr/bin/server'
    environment:
      PORT: 5000
    deploy:
      replicas: 5
      placement:
        constraints:
        - "node.labels.type==cloudlet"
    volumes:
    - type: volume
      source: FooBar
      target: /opt/foobar_data
      volume:
        nocopy: true
    # ...
\end{minted}
\end{listing}

% As previously explained, in the context of Ainur and ExPECA, we use the term \emph{workload} to refer to a collection of containerized, distributed applications deployed on the testbed for a specific amount of time.
% Ainur employ Docker Swarm for this purpose, thus leveraging

Workloads in Ainur are defined using a dictionary of metadata containing a nested \verb|docker-compose| definition, usually provided as a YAML file.
This dictionary is parsed into a \mintinline{python}{WorkloadSpecification} dataclass, which validates the structure of the definition, and which in turn is passed to a \mintinline{python}{DockerSwarm} class instance to actually deploy the workload.

In terms of structure, the workload definition dictionary is expected to contain the following keys:

\begin{description}
    \item[name]
    A string representing an identifier for the workload, used to tag logs and stored data, as well as to identify the generated dynamic distributed storage volumes.
    All leading and trailing whitespace will be trimmed from this value at parse time.
    \item[author]
    An identifier for the author of the workload specification.
    Optional, used in logs.
    \item[email]
    Contact email for the author.
    Optional, used in logs.
    \item[url]
    URL for information related to the workload; optional.
    \item[version]
    String representing a version tag for this workload.
    Optional, but defaults to \texttt{1.0} if not specified, and is parsed into a Python \mintinline{python}{LooseVersion}object for easy manipulation in code.
    \item[max\_duration]
    Represents the maximum duration for the workload, and can be specified as any string value supported by the \texttt{pytimeparse} library\footnote{\url{https://pypi.org/project/pytimeparse/}}.
    For instance, a maximum duration of \SI{1}{\hour} and \SI{30}{\minute} could be specified as any of the following strings: \mintinline{python}{'1h 30m'}, \mintinline{python}{'1.5 h'}, \mintinline{python}{'90m'}, \mintinline{python}{'5400s'}, and so on --- see the \verb|pytimeparse| library documentation for more supported formats.

    This value is used as a timeout for the workload.
    The \mintinline{python}{DockerSwarm} in charge of the execution and orchestration keeps track of the time the workload has been running, forcefully tearing it down if it does not exit cleanly before it the timeout.
    This prevents workloads stuck in non-ending loops from running forever and hogging up testbed resources.

    Optional, defaults to a very liberal \mintinline{python}{'1d'} (i.e.\ \SI{24}{\hour}) if not explicitly specified.
    Note that users are strongly encouraged to explicitly define this value, as the default is subject to arbitrary change as the usage patterns of the testbed evolve.

    \item[compose]

    Corresponds to a nested \texttt{docker-compose v3}\footnote{\url{https://docs.docker.com/compose/compose-file/compose-file-v3/}} definition, and is the single most important component of the workload specification.
    This definition must be compliant with at least version 3 of the \texttt{docker-compose} file specification standard.
    It must additionally be fully compatible with Docker Swarm service stacks\footnote{%
        \url{https://docs.docker.com/engine/reference/commandline/stack_deploy/}%
    }, as the underlying implementation of the \mintinline{python}{DockerSwarm} class uses the \texttt{docker stack deploy} command to deploy the workload on the cluster.
    Each service in this compose specification must contain a \texttt{deploy} configuration segment detailing the scaling and placement requirements for it.
    Finally, if distributed storage has been configured using the components detailed in \cref{sec:storage}, it should be mounted on the appropriate services.
    Examples of these configurations can be seen in \cref{lst:compose}; for more details refer to the \texttt{docker-compose} online documentation.
\end{description}