### 2022-10-25 model.tex
verification

\section{A model of human behavior for \ac{WCA}}\label{sec:model}

In order to accurately emulate the behavior of a human, a model for \ac{WCA} needs to implement two main behaviors.
One, it needs to generate realistic execution times for each step in the task, considering the current and historical impairment of the \ac{WCA} system.
We detail this in \cref{ssec:model:exectimes}.
And two, it needs to produce sequences of input samples for each step mimicking what a real human would generate; this is explained in \cref{ssec:model:frames}.

In the following subsections, \cref{ssec:model:exectimes,ssec:model:frames}, we explain how we design a model which fulfills these requirements.
We briefly discuss where implementations of this model can be obtained in \cref{ssec:model:obtaining}
\todo[inline]{Finish}

\subsection{Generating realistic execution times}\label{ssec:model:exectimes}
\todo[inline]{Rewrite! New model!}

\medskip{}
Two versions of the aforementioned model for the generation of execution times are implemented in Python 3.10.
Apart from a version which follows the above description exactly, we also implement a version which after grouping the data fits an \ac{exGaussian} distribution to each group.
Execution times are then sampled directly from these distributions instead of from the empirical data.
The \ac{exGaussian} distribution was chosen due to the vast body of research supporting its suitability for the modeling of the timing of human actions~\cite{Rohrer1994analysis,Palmer2011shapes,Marmolejo2022generalised}.

\subsubsection{Verifying the behavior of the timing model}\label{ssec:model:verification}
% \todo[inline]{Based on the 4 main conclusions of the PLOS paper.}

We verify the behavior of the above described model with respect to the four main conclusions of our work in~\cite{olguinmunoz:impact2021}, which were previously discussed in \cref{ssec:plos}.
These are reformulated as objectives below:

\begin{enumerate}
    \item\label{it:ttftoexectime} Higher \acp{TTF} should result in higher execution times.
    \item\label{it:duration} When subject to a series of steps at the same level of impairment, desired model behavior depends on the impairment:
    \begin{enumerate}
        \item At low \acp{TTF}, the model should speed up; i.e.\ execution times should decrease.
        \item At medium \acp{TTF}, execution times should remain more or less the same.
        \item At high \acp{TTF}, execution times should progressively increase.
    \end{enumerate}
    \item The effects on execution times due to past reductions or improvements in system responsiveness should linger for at least a few steps whenever system responsiveness changes/
    \item\label{it:neuro} Finally, neuroticism should act as a modulating factor for the above effects.
\end{enumerate}

\begin{figure}
    \centering
    \includegraphics[width=.9\columnwidth]{figs/new_model/ttf_to_exectime.png}
    \caption{%
        Effects of feeding three different \acp{TTF} (\emph{low}, \SI{0}{\second}, \emph{medium}, \SI{2.5}{\second}, or \emph{high}, \SI{5}{\second}) into the model on the generated execution times.
        Higher \acp{TTF} directly lead to higher execution times.
        Error bars indicate the \SI{95}{\percent} \ac{CI}.
    }\label{fig:ttf_to_exectime}
\end{figure}

\cref{fig:ttf_to_exectime} shows the mean execution time outputted by the model when fed three different levels of \ac{TTF} (\emph{low}, \SI{0}{\second}, \emph{medium}, \SI{2.5}{\second}, or \emph{high}, \SI{5}{\second}).
These results were generated by first warming up the model by feeding it \num{25} \acp{TTF} selected at random from the data before feeding it the desired input \ac{TTF}, and recording the generated execution time.
This procedure is repeated \num{600} times for each configuration and target \ac{TTF}.
The resulting mean execution times match precisely the desired behavior mentioned in \cref{it:ttftoexectime}, with the difference in mean execution times at low versus high \acp{TTF} reaching \SI{14}{\percent} (\textasciitilde\SI{5.6}{\second} to \textasciitilde\SI{6.4}{\second}) in the worst case (high neuroticism configuration).
Additionally, we can observe the effects of neuroticism on generated execution times, as specified in \cref{it:neuro}.
At low neuroticism, the average difference between execution times at low versus high \acp{TTF} was of roughly \SI{6.2}{\percent}, compared to \SI{12.5}{\percent} at high neuroticism, clearly putting into evidence the modulating effect of this variable.

\begin{figure}
    \centering
    \includegraphics[width=.9\columnwidth]{figs/new_model/exectime_over_steps.png}
    \caption{%
    Effects of prolonged exposure to constant levels of system impairment on the model.
    At low (\SI{0}{\second}) \ac{TTF}, the models speed-up over time; conversely at high (\SI{5.0}{\second}) \ac{TTF}, the models either present no change or drastically increase their generated execution times, depending on the level of neuroticism.
    Error bars indicate \SI{95}{\percent} \ac{CI}.
    }\label{fig:exectimeduration}
\end{figure}

Next, \cref{fig:exectimeduration} shows the evolution of generated execution times while the model is subject to a fixed \ac{TTF}, either \emph{low} (\SI{0}{\second}) or \emph{high} (\SI{5}{\second}).
These results were generated by first warming up the model with \num{25} random \acp{TTF}, and recording the generated execution times over a sequence of \num{12} steps at a fixed \ac{TTF}; this procedure is repeated \num{600} times for each configuration and target \ac{TTF}.
Once again, we see here behavior matching what is expected of the model, in particular with respect to \cref{it:duration}.
At low \acp{TTF}, the model is on average, across all configurations, \SI{8.2}{\percent} faster at step \num{12} when compared to step \num{1}.
At high \acp{TTF}, the behavior changes depending on the level of neuroticism of the model.
Low neuroticism models basically do not change their execution times, whereas high neuroticism configurations are on average \SI{10}{\percent} slower after \num{12} steps.
This is once again in line with our previous findings, as we had previously concluded that humans tend to speed up during a task, but that this speed-up is hindered and eventually reversed as system responsiveness decreases, and that the strength of this effect is correlated with neuroticism.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figs/new_model/transitions.png}
    \caption{%
        Effects of changes in system impairment on subsequently generated execution times.
        These effects linger on after the change, and thus execution times immediately after a transition are either consistently lower or higher than otherwise at the new \ac{TTF}, depending on the old \ac{TTF}.
        Error bars indicate \SI{95}{\percent} \ac{CI}.
    }\label{fig:transitions}
\end{figure*}

Finally, in \cref{fig:transitions} we showcase the behavior of the model when comparing execution times generated immediately after a change in system responsiveness.
We generate these results by first warming up the model by feeding it a fixed \ac{TTF} (which we will refer to as the \emph{origin} \ac{TTF}) \num{25} times.
Next, another \ac{TTF} value (the \emph{destination} \ac{TTF}) is fed to the model, and we record the output execution time.
Each sample is tagged according to the relation between origin and destination \ac{TTF}, either lower to higher, higher to lower, or equal.
As before, we run \num{600} repetitions of this procedure for each combination of model configuration, origin \ac{TTF}, and destination \ac{TTF}.
Execution times generated immediately after a transition from a higher \ac{TTF} into a lower one are consistently higher than execution times generated without an preceding change in \acp{TTF}.
Conversely, execution times are consistently lower than otherwise immediately after a change from a lower \ac{TTF} into a higher one.
These results are once again in line with our findings in~\cite{olguinmunoz:impact2021}, in which we found lingering effects of transitions between levels of system impairment on human execution times.

\subsection{Generating realistic samples}\label{ssec:model:frames}

Apart from the aforementioned timing and performance data, for~\cite{olguinmunoz:impact2021} we also recorded all collected video frame samples together with matching metadata.
For each video frame submitted to the \ac{WCA} during the tasks, we recorded
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the raw video frame captured
    \item sample submission timestamp
    \item \ac{WCA} processing completed timestamp
    \item result or acknowledgement returned timestamp
    \item a tag representing the result of the \ac{WCA} processing
\end{enumerate*}.
The tags assigned corresponded to:
\begin{description}[font={\bfseries\ttfamily}]
    \item[SUCCESS:] frames which triggered a transition to a new step in the logical task model of the \ac{WCA}, and thus cause the generation of feedback to the user.
        This tag is also used for frames which corrected a previous mistake.
    \item[REPEAT:] frames which contained the same board state as the previous successful frame, and thus produced no feedback.
    \item[LOW\_CONFIDENCE] frames for which the image recognition algorithm in the \ac{WCA} did not reach the necessary confidence threshold to interpret it as a valid board state.
        These frames also produce no feedback.
    \item[BLANK] frames in which not enough of the board is visible due to noise, movement, occlusion, etc.
        These frames produce no feedback either.
    \item[TASK\_ERROR] frames which contained an incorrect board state and thus triggered a transition to a procedural corrective step and the generation of feedback to the user.
        However, it must be noted that none of the \num{40} participants made any mistakes during the task, and thus no such frames were encountered in the data.
\end{description}

We correlate this frame data with the step timing data described in \cref{ssec:model:exectimes} to match frames with their corresponding step execution times.
We assign to each frame a normalized instant value corresponding to its capture instant (expressed in seconds since the start of the step) divided by the total execution time of the step.
To exemplify, for a step with execution time \( t_\text{exec} = \SI{10}{\second} \), a frame captured at time \( \tau = \SI{3}{\second} \) from the start of the step, will have a normalized instant value \( t_\text{norm} \):
\begin{align}
    t_\text{norm} = \frac{\tau}{t_\text{exec}} = \frac{\SI{3}{\second}}{\SI{10}{\second}} = 0.3
\end{align}

This allows us to then analyze the distribution of frame tag probabilities as a step progresses, independently of execution times.
This is illustrated in \cref{fig:frameprobs}.
Intuitively, \texttt{REPEAT} frames dominate the early instants after a step transition, as the user has not had time to start performing the new instruction and thus the \ac{WCA} keeps capturing frames representing the previous state of the board.
As the user starts moving and performing actions, \texttt{BLANK} frames start to dominate, as this activity prevents the \ac{WCA} from capturing ``clean'' frames.
Finally, it must be noted that \texttt{SUCCESS} frames are not included in this probability density plot, as, by definition (\( t_\text{norm} \) represents the normalized instant value):
\begin{equation}
    P(\text{\texttt{SUCCESS}} | t_\text{norm}) =
    \left\{ \begin{array}{ll}
        0 & t_\text{norm} < 1.0    \\
        1 & t_\text{norm} \geq 1.0
    \end{array} \right.
\end{equation}
That is, any frame captured immediately at or after the execution time has been reached will contain the finished board state and thus correspond to a \texttt{SUCCESS} frame.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{model_data/frame_probabilities.png}
    \caption{%
        Probability density of frame result tags as a step progresses.
        Note that \texttt{SUCCESS} frames are not included as --- by definition --- the probability for success frames is \num{1} for all normalized instant values greater than or equal to \num{1.0}.
    }\label{fig:frameprobs}
\end{figure}

Using the above insights, together with the corresponding recorded video frames, we devise an scheme for the procedural generation of a synthetic trace for any step in \ac{WCA} task in the same category as those used in~\cite{olguinmunoz:impact2021}.
We first prepare a discretized representation of the probability density map in \cref{fig:frameprobs}.
We segment the normalized instant value into a number of discrete bins (\num{25} in this work), and calculate the relative fraction of frames for each category in each bin.
For each step, given
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item a collection of random non-\texttt{SUCCESS}, non-\texttt{REPEAT} frames (at least one frame for each of the \texttt{BLANK} and \texttt{LOW\_CONFIDENCE} categories)
    \item an appropriate \texttt{SUCCESS} video frame containing the correct state for the step
    \item an appropriate \texttt{REPEAT} video frame containing the correct state for the \emph{previous} step
\end{enumerate*},
we can then procedurally generate a trace by randomly selecting appropriate frames according to the distributions presented in \cref{fig:frameprobs}.
In other words, for each sampling instant in a step with a given execution time \( t_\text{exec} \):
\begin{enumerate}
    \item We calculate the normalized instant value \( t_\text{norm} = \frac{\tau}{t_\text{exec}} \).
    \item If \( t_\text{norm} \ge 1.0 \), we select the \texttt{SUCCESS} frame and finish the trace.
    \item If instead \( t_\text{norm} < 1.0 \), we find the appropriate bin for \( t_\text{norm} \) and then select a frame by performing a weighted random sampling of the frame categories in the normalized instant bin using the relative proportions for each category.
\end{enumerate}

\subsection{Implementation}
\todo[inline]{Maybe not needed.}

\subsection{Model validation}
\todo[inline]{Implications for task durations, wrt to naive scheme.}

\subsection{Obtaining the model}\label{ssec:model:obtaining}

We provide the model implementations in Python~\num{3.10}, the base data, as well as the client and server implementations used for \cref{ssec:model:verification} to the community as \ac{FOSS}.
All of these are published on the \href{https://github.com/KTH-EXPECA/EdgeDroid2}{\texttt{KTH-EXPECA/EdgeDroid2}} repository on GitHub under a permissive Apache version 2 license.
Furthermore, the client and server implementations are also made available as pre-packaged Docker containers on Docker Hub, under the \href{https://hub.docker.com/r/expeca/edgedroid2}{\texttt{expeca/edgedroid2}} repository.