### 2022-11-13 abstract.tex
pre-shorten draft

\begin{abstract}
    \acf*{WCA} applications present a challenge to benchmark and characterize due to their human-in-the-loop nature.
    Recruiting humans is cumbersome and expensive, and human-subject research is often hard to repeat and replicate.
    Considering the intended mass-scale deployment of these applications in the future, there exists a real need for tools enabling human-independent benchmarking.

    We present in this paper the first data-driven model which allows for the complete end-to-end emulation of humans in \acs*{WCA}.
    We demonstrate the utility of this model by employing it to study application lifetime footprints, where we find an up to \SI{12}{\percent} difference with respect to first order approximations.
    Further, we use this model to develop a novel framework for stochastic optimization of resource consumption-responsiveness tradeoffs in \acs*{WCA}.
    We show that our framework can provide up to a \SI{50}{\percent} reduction in processed frame samples and up to a \SI{20}{\percent} reduction in energy consumption when compared to the state-of-the-art.
\end{abstract}