### 2022-11-18 body/part1/introduction.tex

\chapter{Introduction}

\section{Background \& Related Work}

Deep focus on theoretical background.

\subsection{Theoretical background}
\textcite{Satya2009Case}.
Other papers detailing the advantages of Edge computing.

Next, what applications benefit from the Edge?
\ac{WCA}! Mention both the main body of work (by Satya et.\ al.) but also other authors out there.
Main points: prototypes and characterization of system latencies and load.

\section{Presentation of research questions}

How will human-in-the-loop/\ac{WCA} systems look in production?

How will they look in large scale deployments?

How many applications can given systems support?

How can these dimensions be studied?
Here's where modeling and empirical data-driven research comes in.

Is there potential for optimization?
Yes, trading of system responsiveness for decrease in system load.
Underlying principle under aperiodic, adaptive sampling.
The potential for optimization is novel and stems from \textcite{Olguin:ImpactWCA2021}.