### 2022-12-04 abstract.tex
Satya & Bobby comments

\begin{abstract}
    \acf*{WCA} applications present a challenge to benchmark and characterize due to their human-in-the-loop nature.
    Employing user testing to optimize system parameters is generally not feasible, given the scope of the problem and the number of observations needed for to detect small but important effects in controlled experiments.
    Considering the intended mass-scale deployment of \ac{WCA} applications in the future, there exists a need for tools enabling human-independent benchmarking.

    We present in this paper the first data-driven model which allows for the complete end-to-end emulation of humans in \acs*{WCA}.
    We demonstrate the utility of this model by employing it to study application task durations, where we find an up to \SI{12}{\percent} difference with respect to first order approximations.
    Further, we use this model to develop a novel framework for stochastic optimization of resource consumption-responsiveness tradeoffs in \acs*{WCA}.
    We show that our framework can provide up to a \SI{50}{\percent} reduction in processed frame samples and up to a \SI{20}{\percent} reduction in energy consumption when compared to the state-of-the-art.
\end{abstract}