### 2022-12-19 body/glossary.tex
adding glossary

\newabbreviation{WCA}{WCA}{Wearable Cognitive Assistance}
\newabbreviation{AR}{AR}{Augmented Reality}
\newabbreviation{VR}{VR}{Virtual Reality}
\newabbreviation{XR}{XR}{eXtended Reality}
\newabbreviation{GPU}{GPU}{Graphical Processing Unit}
\newabbreviation{FPS}{FPS}{Frames Per Second}
\newabbreviation{DNN}{DNN}{Deep Neural Network}
\newabbreviation{CPS}{CPS}{Cyber-Physical System}
\newabbreviation{NCS}{NCS}{Networked Control System}
%\acroindefinite{NCS}{an}{a}

\newabbreviation{LAN}{LAN}{Local-Area Network}
\newabbreviation{WLAN}{WLAN}{Wireless Local-Area Network}
\newabbreviation{KPI}{KPI}{Key Performance Indicator}
\newabbreviation{WPAN}{WPAN}{Wireless Personal Area Network}
\newabbreviation{CSMA/CD}{CSMA/CD}{Carrier-Sense Multiple Access with Collision Detection}
\newabbreviation{CLEAVE}{CLEAVE}{ControL bEnchmArkin`g` serVice on the Edge}
\newabbreviation{OS}{OS}{Operating System}
\newabbreviation{UDP}{UDP}{User Datagram Protocol}
\newabbreviation{TCP}{TCP}{Transmission Control Protocol}
\newabbreviation{RMS}{RMS}{Root Mean Square}
\newabbreviation{RTT}{RTT}{Round-Trip Time}
\newabbreviation{CI}{CI}{Confidence Interval}
\newabbreviation{AP}{AP}{Access Point}
\newabbreviation{API}{API}{Application Programming Interface}
\newabbreviation{SSF}{SSF}{Swedish Foundation for Strategic Research}
\newabbreviation{TECoSA}{TECoSA}{Trustworthy Edge Computing Systems and Applications}
\newabbreviation{ABC}{ABC}{Abstract Base Class}
%\acroplural{ABC}[ABCs]{Abstract Base Classes}

\newabbreviation{URL}{URL}{Uniform Record Locator}
\newabbreviation{AWS}{AWS}{Amazon Web Services}
\newabbreviation{EC2}{EC2}{Elastic Compute 2}
\newabbreviation{AMI}{AMI}{Amazon Machine Image}
\newabbreviation{SSH}{SSH}{Secure Shell}
\newabbreviation{IP}{IP}{Internet Protocol}
\newabbreviation{VPN}{VPN}{Virtual Private Network}
\newabbreviation{NAT}{NAT}{Network Address Translation}
\newabbreviation{NTP}{NTP}{Network Time Protocol}
%\acroindefinite{NTP}{an}{a}

\newabbreviation{SDR}{SDR}{Software-Defined Radio}
%\acroindefinite{SDR}{an}{a}

\newabbreviation{VLAN}{VLAN}{Virtual Local-Area Network}
\newabbreviation{APT}{APT}{Advaced Packaging Tool}
\newabbreviation{LTE}{LTE}{Long-Term Evolution}
\newabbreviation{SSID}{SSID}{Service Set Identifier}
\newabbreviation{EPC}{EPC}{Evolved Packet Core}
\newabbreviation{UE}{UE}{User Equipment}
\newabbreviation{eNodeB}{eNodeB}{Evolved Node B}
\newabbreviation{DNS}{DNS}{Domain Name System}
\newabbreviation{MNIST}{MNIST}{Modified National Institute of Standards and Technology}
\newabbreviation{HTTP}{HTTP}{Hyper-Text Transfer Protocol}
%\acroindefinite{HTTP}{an}{a}

\newabbreviation{YAML}{YAML}{YAML Ain't Markup Language}
\newabbreviation{O-RAN}{O-RAN}{Open Radio Access Network}
\newabbreviation{FOSS}{FOSS}{Free and Open Source Software}
\newabbreviation{COSMOS}{COSMOS}{Cloud enhanced Open Software defined MObile wireless testbed for city-Scale deployment}
\newabbreviation{OMF}{OMF}{ORBIT Management Framework}
\newabbreviation{POWDER}{POWDER}{Platform for Open Wireless Data-driven Experimental Research}
\newabbreviation{COTS}{COTS}{Commercial Off-The-Shelves}
\newabbreviation{RF}{RF}{Radio-Frequency}
\newabbreviation{VM}{VM}{Virtual Machine}
\newabbreviation{REST}{REST}{REpresentational Sate Transfer}
\newabbreviation{MaaS}{MaaS}{Metal-as-a-Service}
\newabbreviation{OEDL}{OEDL}{\ac{OMF} Experiment Description Language}
\newabbreviation{RSpec}{RSpec}{Resource Specification}
\newabbreviation{LXC}{LXC}{LinuX Containers}
\newabbreviation{MEC}{MEC}{Mobile Edge Computing}
\newabbreviation{USRP}{USRP}{Universal Software Radio Peripheral}
\newabbreviation{OAI}{OAI}{OpenAirInterface}
\newabbreviation{NR}{NR}{New Radio}
\newabbreviation{FPGA}{FPGA}{Field-Programmable Gate Array}
\newabbreviation{ARA}{ARA}{Agricultural and ruRAl communities}
\newabbreviation{UHF}{UHF}{Ultra-High Frequency}
\newabbreviation{MIMO}{MIMO}{Multi-Input Multi-Output}
\newabbreviation{CHI}{CHI}{CHameleon Infrastructure}
\newabbreviation{CMU}{CMU}{Carnegie Mellon University}
\newabbreviation{GSR}{GSR}{Galvanic Skin Response}
\newabbreviation{EEG}{EEG}{electroencephalography}
\newabbreviation{BFI}{BFI}{Big Five Inventory of personality traits~\cite{john1999big}}
\newabbreviation{ITQ}{ITQ}{Immersive Tendencies Questionnaire~\cite{witmer1998measuring}}