### 2022-12-19 body/part1/futurework.tex
glossaries

\chapter{Open questions \& future work}

Extending methodology to other classes of \gls{WCA}.
Ping-pong is the typical example --- look for other classes too.

Extending methodology to other systems.
\gls{AR}, \gls{XR} in general?
Closed-loop control? (Although we already started with CLEAVE).

Further optimization of \gls{WCA} systems.
In this thesis we touch upon sampling, but other dimensions can also be optimized.
Resource allocation, scheduling, user identification and classification, QoS/QoE guarantees?