### 2022-12-19 publications/2022CLEAVE/acks.tex

\section*{Acknowledgements}\label{sec:acks}

This research has been partially funded by
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the VINNOVA Competence Center for \gls{TECoSA} at KTH Royal Institute of Technology
    \item the \gls{SSF}, through grant number ITM17--0246.
\end{enumerate*}

We also wish to thank S.~S.~Mostafavi, V.~N.~Moothedath, M.~Alsakati, O.~Ferm, S.~Vojcic, and A.~Bhattacharya for their support and valuable contributions to this work.

Finally, we would like to thank the reviewers for their insightful comments and suggestions, which greatly helped us improve this work.