### 2023-01-17 body/part1/overview/implications-optimization.tex

\section{Implications for the optimization of \acs{WCA}}

In this section, we will discuss the implications of the methodology for the optimization of \gls{WCA} applications, in particular when employing the experimental results and the improved model of human timings discussed in \cref{sec:extendingmethodology}.
The version of this discussion presented here is a summarized version of what \cref{paper:olguinmunoz2023realistic} exposes.

\todo[inline]{Do we need more experiments?}

\subsection{Implications with respect to application lifetime footprints}

\subsection{Implications with respect to number of samples per step}

\subsection{Implications with respect to raw energy consumption}