### 2023-01-21 body/part1/overview/implications-optimization.tex
writing

\section{Implications for the optimization of \acs{WCA}}

In this section, we will discuss the implications of the methodology for the optimization of \gls{WCA} applications, in particular when employing the experimental results and the improved model of human timings discussed in \cref{sec:extendingmethodology}.
The discussion presented here is a summarized version of what \cref{paper:olguinmunoz2023realistic} exposes.

\todo[inline]{Do we need more experiments?}

\subsection{Implications with respect to task completion times}

The term \emph{task completion time}, in the context of \gls{WCA}, will refer to the time it takes a user to work their way through the task presented by an assistant application.
In \cref{paper:olguinmunoz2023realistic} we refer interchangeably to task completion times as application lifetimes.
This metric directly relates to system resource and energy consumption --- longer task completion times mean that the system is occupied for longer and thus uses more resources.
It thus presents a relatively straightforward yet interesting starting point for the study of the implications of the model on the optimization of \gls{WCA} systems.

\begin{figure}
    \centering
    \includegraphics[height=13em]{Figs/2023EdgeDroid2/task_durations_diff}
    \caption{Difference in task completion times of the realistic model with respect to the reference baseline.
    Bars indicate \SI{95}{\percent} \glspl{CI}.}\label{fig:taskcompletiontimesdiff}
\end{figure}

The main question we aim to ask relates to the consequences of using our model of human timing behaviors versus using a less realistic approximation.
Specifically, we aim to quantify the difference in task completion times when using the realistic model compared to a baseline which does not take into consideration higher order effects on execution times.
In \cref{fig:taskcompletiontimesdiff} we present the results of an initial exploration of this difference, comparing the total task durations of simulated \num{100}-step tasks.
Our reference baseline corresponds here to a first order approximation to empirical execution time modeling, using an \gls{exGaussian} distribution fitted to \emph{all} the data points from \cref{paper:olguinmunoz2021impact} (without any grouping whatsoever) which is randomly sampled to obtain execution time values.

These results show clear and significant differences in task durations between the realistic model and the baseline, with a reduction of up to \SI{11}{\percent} at low \glspl{TTF} for both levels of neuroticism.
At higher \glspl{TTF}, the high neuroticism parameterization also results in task durations more than \SI{11}{\percent} longer than the baseline.

\subsection{Implications with respect to number of samples per step}

\subsection{Implications with respect to raw energy consumption}