### 2023-01-25 body/part1/overview/implications-optimization.tex

\section{Implications for the optimization of \glsfmtshort{WCA}}