### 2023-02-14 body/part1/futurework.tex
futwork

\section{Conclusions}

\section{Open questions \& future work}

The present thesis presents a comprehensive exploration into the application of our methodology for \glspl{NCS} and specially for \gls{WCA}.
Nonetheless, a number of key questions remain open for future research in this field.

The first concern is the applicability of the methodology to other systems and applications in the context of Edge Computing.
This thesis explores the methodology's use for step-based \gls{WCA} in depth, and also performs a less thorough case study on its benefits for \gls{NCS} research.
The question remains if the methodology is suitable for
\begin{inlineenum}
    \item other, non-\gls{WCA} \gls{XR} applications, such as more general mobile \gls{AR} and \gls{VR}
    \item other more general categories of \glspl{CPS}
\end{inlineenum}.
The results obtained thus far indicate that the methodology is applicable to a wide range of interactive and \gls{CPS} applications in Edge Computing, but these results should be confirmed by domain-specific studies.

In accordance with the broader questions mentioned above, this thesis also raises specific questions related to its individual contributions to the literature.
The EdgeDroid tool, introduced in papers \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid,paper:olguinmunoz2023realistic}, has the potential for further development to support other forms of \gls{WCA} applications.
Currently, the tool is only designed for a specific class of step-based \gls{WCA} with a linear task model.
However, the tool, and by extension the methodology, could be easily adapted to similar applications.

We identify the potential for expanding the dataset presented in \cref{paper:olguinmunoz2021impact} to include a more diverse range of participants.
The current dataset is limited to young undergraduate students from a highly selective university in the \gls{USA}.
Adding participants from different demographic groups would enhance the representativeness of the data and increase the generalizability of the findings.
This would also provide an opportunity to extend the human timing model presented in \cref{paper:olguinmunoz2023realistic} to better reflect the behavior of a broader range of users.

The optimization framework described in \cref{paper:olguinmunoz2023realistic} can be enhanced in two ways.
Firstly, the framework assumes that human execution times follow a Rayleigh distribution, but in reality, they are often modeled with an \gls{exGaussian} distribution.
Secondly, the framework can be made more practical by incorporating the optimization of scenarios with multiple clients and a distributed solution.

There are numerous avenues for further investigation in the optimization and development of \gls{WCA}.
While this thesis has explored the optimization of sampling and energy consumption, there are several other dimensions of these systems that can be examined.
These include resource allocation, task scheduling, user identification and classification, as well as guaranteeing both \gls{QoE} and \gls{QoS}.
Each of these areas presents opportunities for future work and could lead to significant advancements in the field of \gls{WCA}.

\medskip
Our research on applying the methodology to \glspl{NCS} presents promising potential for future work.
While \cref{paper:olguinmunoz2022cleave} provides an initial exploration, further investigation through more realistic and thorough experimentation is necessary to validate the results obtained by the framework.
To this end, we aim to conduct larger-scale experiments that utilize cutting-edge networking technologies, including 5G and future advancements such as 6G.
By exploring the full range of networking technologies and scenarios, we can ensure that our methodology is robust and applicable to a wide range of \gls{NCS}.

We also intend to extend the functionality of the \gls{CLEAVE} tool significantly, primarily to target a wider variety of \glspl{NCS} deployments.
Currently, \gls{CLEAVE} includes a small repertoire of emulations and controllers.
We are planning to extend this with the goal of creating an open library of \glspl{NCS} to share with the community.

\medskip
Finally, we see tremendous opportunities for testbed-driven research in Edge Computing.
Our current efforts in this context aim towards the extension of our Ainur framework and its integration with \gls{CHI}~\cite{keahey2020lessons}.
We believe Ainur's unique focus on fully-automated end-to-end wireless edge computing experimentation makes it valuable addition to the \gls{CHI} stack.
At the same time, integrating with \gls{CHI} will provide a number of features and functionalities which will greatly expand Ainurs potential, such as resource reservation, multi-tenant testbed access, and automated bare-metal and \gls{VM}-host instantiation.

The \gls{EXPECA} testbed is undergoing a major expansion and improvement.
%The aim is to enhance its capabilities and hardware to meet the ever-increasing demands of the industry.
The expansion consists of three main components;
\begin{inlineenum}
    \item a hardware upgrade
    \item a small scale, private 5G deployment
    \item a campus-wide 5G deployment
\end{inlineenum}.
Firstly, the \gls{EXPECA}s hardware is receiving a complete overhaul.
We are replacing our initial Raspberry Pi-based prototype with a number of industry grade servers.
Secondly, a small-scale private 5G deployment using industrial-grade equipment is currently being deployed.
This will provide a controlled environment for testing and evaluating 5G technologies and applications.
The use of industrial-grade equipment ensures that the results obtained from the tests are reliable and can be directly applied to real-world scenarios.
Thirdly, the testbed will be extended to a campus-wide deployment at \gls{KTH} in Stockholm.
This will provide an opportunity for a larger-scale deployment of 5G and \gls{MEC} technologies and applications, allowing for testing in real-world settings.
The campus-wide deployment will also serve as a platform for researchers and industry professionals to collaborate and exchange ideas.
The combined efforts of the small-scale private deployment and the campus-wide deployment at \gls{KTH} will significantly enhance the capabilities and knowledge of the \gls{EXPECA} testbed, positioning it as a leading platform for 5G and \gls{MEC} research and innovation.

In conclusion, the present dissertation provides a comprehensive examination of the methodology for \glspl{NCS} and \gls{WCA}, but there are still many open questions that need to be addressed in future research.
The EdgeDroid tool and the dataset have the potential to be extended, and the optimization framework employed in the thesis could be improved.
We also plan to extend the functionality of the \gls{CLEAVE} tool and perform large-scale experimentation on novel networking technologies.
There are also significant opportunities for testbed-driven research in Edge Computing through the \gls{EXPECA} testbed.