### 2023-03-04 publications/2023EdgeDroid2/acks.tex

\section*{Acknowledgements}\label{sec:acks}

This work has been partially funded by the \gls{SSF} (grant number \verb|ITM17-0246| (ExPECA)), and the \gls{NSF} (grant number \verb|CNS-2106862|).
The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
Any opinions, findings, conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the view(s) of their employers or funding sources.