### 2023-03-09 recyclebin/experimental_research_edge_computing.tex
relatedwork

\subsubsection{Experimental research in Edge Computing}
\todo[inline]{Should be moved to related work or removed.}