### 2023-03-24 body/part1/chapter2-relatedwork/relatedwork.tex
more tweaks

\section{Related work}

\subsection{Emulation in \glsfmtshort{XR} research}

Only a limited number of works addressing the partial or complete emulation of XR applications for benchmarking and research purposes exist in literature.
One such work is~\cite{williams2013transform}.
In this, the authors introduced a trace-based toolkit for the evaluation of tracking algorithms in \gls{AR} research and development.
The trace consists of geographical coordinates and accelerometer data as well as video frames which can be replayed on a desktop visualization tool or a smartphone application for the evaluation of tracking algorithms in \gls{AR} research and development.

\cite{munro2016aaremu} presents a trace-based framework that allows a previously recorded trace of inputs to be replayed to the processing components of a real \gls{AR} application by mocking Android \gls{SDK} \glspl{API}.
This approach allows for benchmarking and evaluation of algorithms and architectures on real data.
A similar approach was taken by~\cite{choi2022emulating}, but the authors instead provided developers of \gls{AR} with a framework that abstracts input into general classes, allowing for device-agnostic processing.

Finally, the authors of~\cite{chetoui2022arbench} employ \gls{AR} emulation to evaluate hardware rather than the software implementation itself.
Their tool, \emph{ARBench}, features emulated \gls{AR} workloads that mimic the load placed on the system by real applications.
This approach is similar to the one presented in this dissertation, but it exclusively considers on-device \gls{AR} processing.
Furthermore, the benchmarks are deterministic, and no emulation of realistic human behavior is attempted.

\subsection{Edge and Cloud Benchmarking}

This dissertation complements existing work on benchmarking distributed system architectures.
A rich body of literature exists with examples of such approaches for Cloud computing.
For example, the \gls{YCSB}~\cite{cooper2010benchmarking} is a data-intensive benchmark for the profiling of data centers and cloud services.
Workload characterization in datacenters leads to a similar benchmark in~\cite{jia2013characterizing}.
Both these approaches focus heavily on data processing and downlink transmission in cloud computing scenarios.

In~\cite{turner2012cmart}, the authors introduce \emph{C-MART}, a benchmark for modern web applications running on the cloud.
The benchmark utilizes modern web technologies, can be deployed automatically at a large scale, and includes generators that mimic remote clients accessing the system over the internet.\@
\cite{malawski2018benchmarking,back2018using} focus instead on serverless computing, using compute-intensive workloads to benchmark major cloud function providers such as \gls{AWS}, Microsoft Azure, and \gls{GCP}.

A number of works have tackled this challenge in the context of Fog and Edge Computing as well.
\emph{EdgeBench}, introduced in~\cite{das2018edgebench}, corresponds to a benchmarking suite for the Edge.
It emulates three different workloads, speech-to-text, image recognition, and sensor network application traffic, and specifically targets serverless deployments.\@
\cite{baurle2022comb} presents \emph{ComB}, a flexible and extensible, application-oriented benchmarking suite for Edge computing.
Its stated purpose is to assist system designers and application developers to evaluate the suitability of Edge deployments.
The framework is designed to decouple orchestration and workload, allowing for easy extension or replacement of the built-in distributed video-analytics benchmark.

\emph{IoTBench}~\cite{lee2019iotbench} is a benchmarking tool for edge-deployed \gls{IOT} applications.
It includes seven representative workloads, and allows researchers to evaluate the performance of these applications on Edge infrastructure through the analysis of computational demands and resource consumption.\@
\cite{mcchesney2019defog} presents a benchmarking suite for the comparison of deployments on the Edge, Fog, or Cloud named \emph{DeFog}.
\emph{DeFog} includes six latency-critical and bandwidth-intensive, Fog- and Edge-relevant workloads, including real-time video detection and a text-to-speech application.
In~\cite{lockhart2020scission}, the authors introduce a tool, \emph{Scission}, for the benchmarking of \glspl{DNN} partitioned across the user device, edge, and cloud continuum.
The tool can also be used to automatically determine the optimal partition of a \gls{DNN}.
Finally, \emph{OpenRTiST}~\cite{george2020openrtist} is a tool for the benchmarking and performance evaluation of Edge Computing deployments using real-time processing of a live video feed.
Video is streamed from a source Android device to a compute node where neural style-transfer~\cite{gatys2016image} is applied to it.
The transformed video is then fed back to the source to be presented, resulting in a compute-intensive, bandwidth-hungry, and latency-sensitive benchmark.

\subsection{Experimental research in \glsfmtshortpl{NCS}}

A significant amount of work has been dedicated to the modeling and performance characterization of \glspl{NCS} due to their potential benefits for industrial and commercial settings~\cite{lu2016real,hespanha2007survey,zhang2013network,zhang2016survey}.
Research in \glspl{NCS} has mostly focused on addressing the challenges imposed by the introduction of best-effort communication networks into traditional control systems.
Network delays (latency), losses, and \emph{jitter} --- referring to the variation in random network delays --- are at the forefront of these challenges.
They have partially been addressed through the clever placement of the controller (e.g.\ by employing Edge Computing)~\cite{sasaki2017layered,sasaki2016vehicle}, as well as through the development of control algorithms capable of accounting for the effects of the network~\cite{zhang2013network}.

Most of the large literature concerning \glspl{NCS}, however, follows a theoretical approach, with only small fraction of it dealing with experimental studies~\cite{zhang2019networked}.
\gls{NCS} studies are complex due to the inter-domain nature of these systems, requiring expertise in the fields of communication networks, computing, and control theory.
However, theoretical approaches can only capture network behaviors at a coarse level, and thus a body of works has emerged in later years involving experimental approaches to \gls{NCS} benchmarking.
Within this body of work, we identify three distinct approaches: fully practical, fully simulated, and emulated approaches that combine components of the first two.

In a fully practical approach, the experimental setup uses real systems and hardware to test and validate the performance of the \gls{NCS}.
This approach is straightforward to implement, and can be very useful for evaluating the real-world applicability of the system, but can be costly and time-consuming.
Examples of the application of this approach can be found in studies such as~\cite{drew2005networked,baumann2018evaluating,li2014wireless,cuenca2019periodic}.
These studies implemented wireless and microcontroller-based systems for control and validated them on a physical prototype.

In~\cite{drew2005networked}, the authors present a \gls{NCS} model which considers both random packet delay and loss on both the sensor and actuator sides of the feedback loop.
They validate their design on a physical testbed consisting of inverted pendula controlled from a central computation point over Wi-Fi.
A similar testbed is employed for validation in~\cite{baumann2018evaluating}, in which the authors develop an evaluation methodology for wireless \gls{NCS}.\@
\cite{li2014wireless} implements a wireless microcontroller-based system for vibration control and validates it on a physical prototype comprising a cantilever beam controlled using an \acs{IEEE} 802.15.4 \gls{WPAN} system.\@ \cite{cuenca2019periodic} discusses the design and implementation of a periodic event-triggered sampling and dual-rate control techniques for wireless control, and validate their contributions on a four-rotor autonomous drone platform controlled over WiFi.
\emph{NCSbench}~\cite{zoppi2020ncsbench}, an open-source \gls{NCS} benchmarking platform designed with reproducibility in mind built using the \citetitle{LEGOMindstormsEV3}~\cite{LEGOMindstormsEV3} platform, is another such example.

Completely simulated setups, on the other hand, rely on computer simulations to evaluate the \gls{NCS}'s performance.
This approach is less expensive and faster than the practical approach but may not fully capture the real-world complexities of the system.
Examples of tools and frameworks for the implementation of this approach can be found in works such as~\cite{andersson2005simulation,eyisi2012ncswt}.

A simulated approach is employed in~\cite{du2009novel}, in which the authors develop a novel \emph{Smith} predictor to compensate for varying latencies in wireless \glspl{NCS} and validate it using the \emph{TrueTime}~\cite{henriksson2002truetime} simulator.
In~\cite{chen2015synchronous} the authors validate and evaluate a number of synchronous control strategies for three-motor setups using completely simulated \glspl{NCS} environments.\@
\cite{wu2012application} introduces the concept of \emph{network predictive control} in and uses it to balance a simulated inverted pendulum over a simulated wireless network.\@
\cite{ma2019optimal} proposes an optimal dynamic scheduling strategy that optimizes performance of multi-loop control systems and validate it by simulating a four-loop control system over an \acs{IEEE} 802.15.4 \gls{WPAN}.

Finally, network and hardware-in-the-loop approaches, which aim to strike a balance between realism and efficiency, have also been employed in literature.
Network-in-the-loop refers to experimental setups in which the \emph{network} is instead simulated or emulated.
Such approaches have, for instance, been used to study the effects of the \gls{CSMA/CD} medium access control mechanism on plant stability, as in~\cite{natale2004inverted}.
Hardware-in-the-loop approaches, referring to \gls{NCS} setups in which a real network interacts with a simulated or emulated control system, are particularly prevalent in the field of \emph{smart grid} control.
An example of the application of such an approach can be found in~\cite{wang2020inverter}, in which the authors introduce and validate a novel three-level coordinated control method for photovoltaic inverters.

\subsection{Testbed research in Edge Computing}

The last few years have seen the emergence of a number of small- to mid-scale platforms and testbeds for Edge Computing research.
Several of these are additionally driven by research interests in novel mobile networking technologies and \gls{MEC}, as well as in the virtualization of network functions.
Of particular interest to us are the
\begin{inlineenum}
    \item \acs{GENI}
    \item Chameleon
    \item \acs{COSMOS}
    \item \acs{POWDER}
    \item \acs{ARA}
    \item EdgeNet
%    \item Drexel Grid \gls{SDR}
\end{inlineenum} testbeds.

\gls{GENI} is a testbed deployed across more than fifty sites across the \gls{USA}.
Originally designed for ``the development, deployment, and validation of transformative, at-scale concepts in network science, services, and security''~\cite{berman2014geni}, \gls{GENI} today combines local edge compute resources with mobile network last-hop connectivity for \gls{MEC} experimentation~\cite{gosain2017geni}.
The wireless-capable sites incorporate single-hop access to compute, storage, and network resource, making \gls{GENI} a compelling Edge Computing platform.

The Chameleon testbed is a research platform originally designed for the study and evaluation of cloud computing systems~\cite{keahey2020lessons}, but which has recently been expanded to target Edge Computing research through the \acs{CHI}@Edge~\cite{chiatedge} project.
Chameleon is a collaborative project operated by a consortium of research institutions and industry partners, and is designed to provide a flexible and reconfigurable infrastructure for researchers to experiment with and test various cloud computing technologies, algorithms, and configurations.
The testbed provides a highly reconfigurable array of hardware and software resources, including servers, storage systems, and networking equipment, able to simulate various cloud and Edge Computing scenarios.

\gls{COSMOS} is a testbed deployed in New York City (State of New York, \gls{USA}), containing over \num{200} rooftop, intermediate, and mobile nodes.
The testbeds main \emph{raison d'être} is \gls{MEC} research, with a strong emphasis on wireless networks.
The system includes \glspl{SDR}, \si{\milli\meter}-wave equipment, optical fibers, cloud integration, and edge compute for core network functionality and application data processing~\cite{yu2019cosmos,raychaudhuri2020challenge}.

\gls{POWDER} is described by the authors of~\cite{breen2020powder} as a ``city-scale, remotely accessible, end-to-end software defined platform supporting a broad range of wireless and mobile related research''.
It features roughly \num{15} fixed programmable radio nodes, based on off-the-shelves \glspl{SDR}, distributed across a \SI{15}{\kilo\meter\squared} area in Salt Lake City (State of Utah, \gls{USA}).
These radio nodes are combined with edge-based compute nodes as well as cloud resources for full \gls{MEC} experimentation.

The \gls{ARA} platform~\cite{zhang2022ara} is a highly flexible testbed based on the \gls{CHI} software suite originally developed for the aforementioned Chameleon testbed~\cite{keahey2020lessons}.
\gls{ARA} is developed by researchers at Iowa State University in Ames, Iowa (\gls{USA}), and spans a rural area with a diameter of over \SI{60}{\kilo\meter}~\cite{zhang2022ara}.
Its core goal is the study and deployment of advanced Edge Computing and wireless platforms and technologies in real-world agricultural and rural settings, and includes a broad range of wireless technologies deployed through both \glspl{SDR} and programmable \gls{COTS} radios, as well as automated ground vehicles, cameras and sensors.

EdgeNet deviates from the approaches described above and represents a fully software-based, massively distributed public platform for Edge Computing research~\cite{cappos2018edgenet,senel2021edgenet1,senel2021edgenet2}.
It consists of a public, modified and customized Kubernetes~\cite{kubernetes} cluster.
The EdgeNet software is available for anyone to download;
individuals and organizations can quickly deploy self-hosted edge nodes which are then made available to anyone using the platform.
The end result is a flexible and scalable global edge cloud simplifying research and prototype development in edge and cloud computing.
A similar, closely-related project is KubeEdge~\cite{xiong2018extend}, a software framework for extending Kubernetes clusters from the cloud to the edge.

Finally, a number of other, small-scale approaches also exist.
In~\cite{gedawy2016cumulus}, the authors propose Cumulus, a prototype testbed for edge compute offloading in \gls{IOT} with the explicit goal of providing a generic and heterogeneous platform for Edge Computing research.
The authors of~\cite{rimal2018experimental} propose an experimental testbed for two-tiered edge architectures in the context of \gls{FIWI} Edge Computing.
This testbed is then employed to validate wireless access and compute resource scheduling strategies in these systems.
In~\cite{yamanaka2021design}, the authors present an experimental Edge Computing testbed capable of determining the offload location of a workload based on desired latency parameters.
\cite{diao2019scalable} propose yet another Kubernetes-based testbed based for edge task offloading based on heterogeneous hardware including low-power single-board computers such as Raspberry Pis and Jetson Nanos.
Finally,\ \cite{moorthy2022cloudraft} proposes \gls{CLOUDRAFT}, a cloud-based framework for mobile network experimentation, with a focus on simplifying the management of testbed resources.
The goal of this project is to integrate, coordinate, share, and improve upon existing testbeds through a common interface.