### 2023-03-28 body/part1/chapter1-introduction/tables/cloud-vs-edge.tex

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage{graphicx}
\begin{table}[]
    \centering
    \caption{Comparison between Cloud Computing and three implementations of Edge Computing.}\label{tab:cloud-vs-edge}
    \tiny
    \renewcommand{\arraystretch}{1.5}
    \begin{tabularx}{\textwidth}{@{}lYYYY@{}}
        \toprule
        &
        \textbf{Cloud} &
        \textbf{Cloudlet} &
        \textbf{\acs{MEC}} &
        \textbf{Fog} \\ \midrule
        \textbf{Location} &
        Data centers that are geographically distant from end-users. &
        Micro-datacenters close to end-users. &
        Integrated within the \acs{RAN}. &
        Distributed across the Edge-Cloud network continuum. \\
        \textbf{Latency} &
        \ensuremath{> \SI{50}{\milli\second}} &
        \ensuremath{\leq \SI{50}{\milli\second}} &
        \ensuremath{\leq \SI{10}{\milli\second}} &
        \ensuremath{\leq \SI{50}{\milli\second}} \\
        \textbf{Resources} &
        Limited only by network bandwidth and latency. &
        \multicolumn{3}{%
            >{\hsize=\dimexpr3\hsize+4\tabcolsep+2\arrayrulewidth\relax}Y%
        }{Limited, but orders of magnitude more than end-user devices.} \\
        \textbf{Security} &
        Relies on strong network and perimeter defenses. &
        \multicolumn{3}{%
            >{\hsize=\dimexpr3\hsize+4\tabcolsep+2\arrayrulewidth\relax}Y%
        }{Data can be kept close to the source and encrypted.} \\
        \textbf{Privacy} &
        Data is aggregated and stored centrally. &
        \multicolumn{3}{%
            >{\hsize=\dimexpr3\hsize+4\tabcolsep+2\arrayrulewidth\relax}Y%
        }{Data can be kept locally, reducing the risk of exposure.} \\
        \textbf{Scalability} &
        Easily scalable by adding more resources to datacenters. &
        \multicolumn{3}{%
            >{\hsize=\dimexpr3\hsize+4\tabcolsep+2\arrayrulewidth\relax}Y%
        }{Limited by virtue of Edge nodes being distributed and energy-constrained in comparison to datacenters.} \\
        \textbf{Use~Cases} &
        Large-scale systems that are not latency-sensitive or do not require real-time processing. &
        Latency-sensitive applications such as \acs{CPS} and \acs{AR}. &
        Latency-sensitive applications in mobile and wearable contexts. &
        \acs{IOT} \\ \bottomrule
    \end{tabularx}%
\end{table}