### 2023-04-03 body/part1/chapter2-relatedwork/tables/scope.tex
scope

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[]
    \centering
    \caption{Overview of the scope and main contribution of each included publication.}\label{tab:scope}
    \tiny
    \renewcommand{\arraystretch}{1.5}
    \begin{tabularx}{\textwidth}{@{}ccXYX@{}}
        \toprule
        \textbf{Paper} & \textbf{Application} & \textbf{Emulation Approach} & \textbf{Network} & \textbf{Main Contribution}                                \\ \midrule
        \ref{paper:olguinmunoz2018demoscaling} & \acs{WCA} & Trace, \acs{FSM} user model & Wi-Fi & Initial demo of methodology. \\
        %
        \ref{paper:olguinmunoz2019edgedroid} & \acs{WCA} & Trace, \acs{FSM} user model & Wi-Fi & EdgeDroid tool for application of the methodology to \acs{WCA}. \\
        %
        \ref{paper:olguinmunoz2021impact} & \acs{WCA} & \acs{N/A} & No network & Characterization of human behavior in \acs{WCA}. \\
        %
        \ref{paper:olguinmunoz2022cleave} & \acs{NCS} & Software implementation of Control System & Wi-Fi & \acs{CLEAVE} tool for application of the methodology to \acs{NCS}. \\
        %
        \ref{paper:olguinmunoz2022ainur} & \acs{WCA}, \acs{NCS} & \acs{N/A} & Ethernet, Wi-Fi, 4G~\acs{LTE} & Ainur framework for Edge Computing testbed automation. \\
        \ref{paper:olguinmunoz2023realistic} &
        \acs{WCA} &
        Data-driven stochastic synthetic trace and user model &
        Wi-Fi &
        Model for end-to-end emulation of humans in \acs{WCA}, exploration of implications of methodology and model for \acs{WCA} optimization. \\
        \bottomrule
    \end{tabularx}
\end{table}