### 2023-04-09 body/part1/chapter2-relatedwork/scope.tex
cps

\section{Scope of the thesis}

\todo[inline]{Write}

\todo[inline]{
    Emphasize that apps are standing for a new class of application class that the community is interested in developing.
    Why?
    Utility.
    Weave into intro and scope.

    Not just a random mashup.
    Relevance of the combination!

    Prime reason for experimental evaluation.
    Capture effects of compute, network, and design of the application.
    Hard to capture otherwise.

    Build argument in state of the art that there is no current satisfying way to address this.
    Potentially orders of magnitude off.

    In scope: interested in meaningful performance evaluation (i.e.\ capture all the complexity!) Make sure we note this.
    Show how apps and system are intimately related to each other and complex to evaluate.
    Where do complexities arise.
    Understanding system performance is key and complex.

    Scope: define the challenges.
}

\begin{enumerate}
    \item Performance evaluation.
    \item \gls{CPS} and \gls{MAR}
    \item Edge Computing 
\end{enumerate}