### 2023-04-10 resources/glossary.tex

%! suppress = MissingImport
\newacronym[longplural={Wearable Cognitive Assistants}]{WCA}{WCA}{Wearable Cognitive Assistance}
\newacronym[longplural={Augmented Realities}]{AR}{AR}{Augmented Reality}
\newacronym[longplural={Virtual Realities}]{VR}{VR}{Virtual Reality}
\newacronym[longplural={eXtended Realities}]{XR}{XR}{eXtended Reality}
\newacronym[longplural={Mixed Realities}]{MR}{MR}{Mixed Reality}
\newacronym{MAR}{MAR}{Mobile \acs{AR}}
\newacronym{GPU}{GPU}{Graphical Processing Unit}
\newacronym{FPS}{FPS}{Frames Per Second}
\newacronym{DNN}{DNN}{Deep Neural Network}
\newacronym{CPS}{CPS}{Cyber-Physical System}
\newacronym[prefixfirst={a\ }, prefix={an\ }]{NCS}{NCS}{Networked Control System}
\newacronym{LAN}{LAN}{Local-Area Network}
\newacronym{WLAN}{WLAN}{Wireless Local-Area Network}
\newacronym{KPI}{KPI}{Key Performance Indicator}
\newacronym{WPAN}{WPAN}{Wireless Personal Area Network}
\newacronym{CSMA/CD}{CSMA/CD}{Carrier-Sense Multiple Access with Collision Detection}
\newacronym{OS}{OS}{Operating System}
\newacronym{UDP}{UDP}{User Datagram Protocol}
\newacronym{TCP}{TCP}{Transmission Control Protocol}
\newacronym{RMS}{RMS}{Root Mean Square}
\newacronym{RTT}{RTT}{Round-Trip Time}
\newacronym{CI}{CI}{Confidence Interval}
\newacronym{AP}{AP}{Access Point}
\newacronym{API}{API}{Application Programming Interface}
\newacronym[user1={Stiftelsen för Strategisk Forskning}]{SSF}{SSF}{Swedish Foundation for Strategic Research}
\newacronym{TECoSA}{TECoSA}{Trustworthy Edge Computing Systems and Applications}
\newacronym[longplural={Abstract Base Classes}]{ABC}{ABC}{Abstract Base Class}
%\acroplural{ABC}[ABCs]{Abstract Base Classes}
\newacronym{URL}{URL}{Uniform Record Locator}
\newacronym{AWS}{AWS}{Amazon Web Services}
\newacronym{GCP}{GCP}{Google Cloud Platform}
\newacronym{EC2}{EC2}{Elastic Compute 2}
\newacronym{AMI}{AMI}{Amazon Machine Image}
\newacronym{SSH}{SSH}{Secure Shell}
\newacronym{IP}{IP}{Internet Protocol}
\newacronym{VPN}{VPN}{Virtual Private Network}
\newacronym{NAT}{NAT}{Network Address Translation}
\newacronym[prefixfirst={a\ }, prefix={an\ }]{NTP}{NTP}{Network Time Protocol}
%\acroindefinite{NTP}{an}{a}
\newacronym[prefixfirst={a\ }, prefix={an\ }]{SDR}{SDR}{Software-Defined Radio}
%\acroindefinite{SDR}{an}{a}
\newacronym{VLAN}{VLAN}{Virtual Local-Area Network}
\newacronym{APT}{APT}{Advaced Packaging Tool}
\newacronym{LTE}{LTE}{Long-Term Evolution}
\newacronym{SSID}{SSID}{Service Set Identifier}
\newacronym{EPC}{EPC}{Evolved Packet Core}
\newacronym{UE}{UE}{User Equipment}
\newacronym{eNodeB}{eNodeB}{Evolved Node B}
\newacronym{DNS}{DNS}{Domain Name System}
\newacronym{MNIST}{MNIST}{Modified National Institute of Standards and Technology}
\newacronym[prefixfirst={a\ }, prefix={an\ }]{HTTP}{HTTP}{Hyper-Text Transfer Protocol}
%\acroindefinite{HTTP}{an}{a}
\newacronym{YAML}{YAML}{YAML Ain't Markup Language}
\newacronym{RAN}{RAN}{Radio Access Network}
\newacronym{O-RAN}{O-RAN}{Open \acs{RAN}}
\newacronym{FOSS}{FOSS}{Free and Open Source Software}
\newacronym{COSMOS}{COSMOS}{Cloud enhanced Open Software defined MObile wireless testbed for city-Scale deployment}
\newacronym{OMF}{OMF}{ORBIT Management Framework}
\newacronym{POWDER}{POWDER}{Platform for Open Wireless Data-driven Experimental Research}
\newacronym{COTS}{COTS}{Commercial Off-The-Shelves}
\newacronym{RF}{RF}{Radio-Frequency}
\newacronym{VM}{VM}{Virtual Machine}
\newacronym{REST}{REST}{REpresentational Sate Transfer}
\newacronym{MaaS}{MaaS}{Metal-as-a-Service}
\newacronym{OEDL}{OEDL}{\acs{OMF} Experiment Description Language}
\newacronym{RSpec}{RSpec}{Resource Specification}
\newacronym{LXC}{LXC}{LinuX Containers}
\newacronym{MEC}{MEC}{Mobile Edge Computing}
\newacronym{USRP}{USRP}{Universal Software Radio Peripheral}
\newacronym{OAI}{OAI}{OpenAirInterface}
\newacronym{NR}{NR}{New Radio}
\newacronym{FPGA}{FPGA}{Field-Programmable Gate Array}
\newacronym{ARA}{ARA}{Agricultural and ruRAl communities}
\newacronym{UHF}{UHF}{Ultra-High Frequency}
\newacronym{MIMO}{MIMO}{Multi-Input Multi-Output}
\newacronym{CHI}{CHI}{CHameleon Infrastructure}
\newacronym{CMU}{CMU}{Carnegie Mellon University}
\newacronym{GSR}{GSR}{Galvanic Skin Response}
\newacronym{EEG}{EEG}{ElectroEncephaloGraphy}
\newacronym{BFI}{BFI}{Big Five Inventory of personality traits}
\newacronym{ITQ}{ITQ}{Immersive Tendencies Questionnaire}
\newacronym{GPS}{GPS}{Global Positioning System}
\newacronym{FSM}{FSM}{Finite State Machine}
\newacronym{CLEAVE}{CLEAVE}{ControL bEnchmArking serVice on the Edge}
\newacronym[longplural={Times-to-Feedback}]{TTF}{TTF}{Time-to-Feedback}
\newacronym{QoE}{QoE}{Quality of Experience}
\newacronym{QoS}{QoS}{Quality of Service}
\newacronym{PCA}{PCA}{Principal Component Analysis}
\newacronym{exGaussian}{Ex-Gaussian}{Exponentially Modified Gaussian}
\newacronym{IEEE}{IEEE}{Institute of Electrical and Electronics Engineers}
\newacronym{CPU}{CPU}{Central Processing Unit}
\newacronym{N/A}{N/A}{Not Applicable}
\newacronym{CIO}{CIO}{Confidence Interval}
\newacronym{MLE}{MLE}{Maximum Likelihood Estimation}
\newacronym{CDF}{CDF}{Cumulative Distribution Function}
\newacronym{CCDF}{CCDF}{Complementary \acs{CDF}}
\newacronym{PDF}{PDF}{Probability Density Function}
\newacronym{ECDF}{ECDF}{Empirical \acs{CDF}}
\newacronym{NSF}{NSF}{United States National Science Foundation}
\newacronym{ANOVA}{ANOVA}{ANalysis Of VAriance}
\newacronym{CV}{CV}{Computer Vision}
\newacronym{SEM}{SEM}{Standard Error of the Mean}
\newacronym{SD}{SD}{Standard Deviation}
\newacronym{BPM}{BPM}{Beats Per Minute}
\newacronym{HR}{HR}{Heart Rate}
\newacronym{IOT}{IoT}{Internet of Things}
\newacronym{ML}{ML}{Machine Learning}
\newacronym{MPC}{MPC}{Model-Predictive Control}
\newacronym{USA}{USA}{United States of America}
\newacronym{CLOUDRAFT}{CloudRAFT}{Cloud-based Remote Access For wireless Testbeds.}
\newacronym{FIWI}{FiWi}{Fiber-Wireless}
\newacronym{GENI}{GENI}{Global Environment for Networking Innovation}
\newacronym{SBC}{SBC}{Single-Board Computer}
\newacronym{EXPECA}{ExPECA}{Experimental Platform for Edge Computing Applications}
\newacronym[user1={Kungliga Tekniska Högskolan}, prefixfirst={the\ }]{KTH}{KTH}{Royal Institute of Technology of Sweden}
\newacronym{HCI}{HCI}{Human-Computer Interaction}
\newacronym{2D}{2-D}{two-dimensional}
\newacronym{3D}{3-D}{three-dimensional}
\newacronym{PD}{PD}{Proportional-Differential}
\newacronym{PID}{PID}{Proportional-Integral-Differential}
\newacronym{SDK}{SDK}{Software Development Kit}
\newacronym{YCSB}{YCSB}{Yahoo! Cloud Serving Benchmark}
\newacronym{ITS}{ITS}{Intelligent Transportation System}
\newacronym{OOS}{OoS}{Out-of-Scope}
\newacronym{OMNET}{OMNeT++}{Objective Modular Network Testbed in C++}
\newacronym{ns2}{ns-2}{network simulator 2}
\newacronym{ns3}{ns-3}{network simulator 3}
\newacronym{IRB}{IRB}{Institutional Review Board}
%\glsaddall%