### 2023-04-23 body/part1/chapter2-relatedwork/tables/scope.tex
scope

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[]
    \centering
    \caption{Visualization of the scope of this dissertation with respect to performance evaluation.
    We focus exclusively on Edge Computing based on Cloudlets~\cite{satyanarayanan2009case} and on experimental performance evaluation of \gls{XR} and \glspl{CPS}.}
    \label{tab:scope}
    \begin{tabularx}{\textwidth}{@{}YYYY@{}}
        \toprule
        & {Analytical Modeling} & {Simulations} & {Experimental}                   \\ \midrule
        Cloudlets & \cellcolor[HTML]{9AFF99}Yes & \cellcolor[HTML]{9AFF99}Yes & \cellcolor[HTML]{9AFF99}{\color[HTML]{333333} Yes} \\
        \gls{MEC} & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{FFCCC9}No                         \\
        Fog & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{FFCCC9}No                         \\
        \gls{XR}  & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{9AFF99}{\color[HTML]{333333} Yes} \\
        \glspl{CPS} & \cellcolor[HTML]{FFCCC9}No              & \cellcolor[HTML]{FFCCC9}No      & \cellcolor[HTML]{9AFF99}{\color[HTML]{333333} Yes} \\ \bottomrule
    \end{tabularx}
\end{table}