### 2023-05-10 body/acks.tex
acks

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ACKNOWLEDGMENT %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Acknowledgements}
\glsresetall%

I would like to begin by thanking my supervisor, Prof. James Gross, for his guidance and patience during this long process.\todo{todo}

I would like to express my deepest gratitude to Prof. Mahadev ``Satya'' Satyanarayanan, for hosting me repeatedly, over long periods of time, at \gls{CMU} as a visiting researcher.
His tremendous experience and generosity have been invaluable, and his guidance and advise have certainly shaped my work.
Likewise, I'm deeply indebted to Prof. Roberta ``Bobby'' Klatzky of \gls{CMU} for the many helpful and insightful discussions, for introducing me to the field of Human-Computer Interaction, and for always greeting me with a warm smile.

I would also like to thank everybody involved in the defense of this thesis.
I am very grateful to Prof. Yu Xiao of Aalto University, for agreeing to act as my dissertation opponent.
My dissertation defence will certainly benefit greatly from her expertise and insights.
I would like to extend my sincere thanks to my grading committe, consisting of Prof. Ana Aguiar (Universidade do Porto), Prof. Per Gunningberg (Uppsala Universitet), and Prof. Klaus Wehrle (\gls{RWTH}), for investing time and effort into reading and grading this dissertation.
Special thanks also go to Prof. Markus Flierl for advance-reviewing this dissertation, and to Prof. Mats Bengtsson for agreeing to act as chairperson during the defense.

I also extend my thanks to my colleagues and co-authors.
To Samie, Vishnu, and Neel at \pgls{KTH}, for always being willing to engage in interesting discussions, but also for generally being fun to be around.
And to Junjue and Babu Pillai from \gls{CMU} for enriching my research with their interesting insights.

I am very lucky and grateful that I got to stay at both \pgls{KTH} and \gls{CMU} over the years, and I would like to thank everyone who made these spaces welcoming and engaging.
A big thank you goes out to my colleagues (both current and former) at the division of \gls{ISE} for all those fun lunches at Restaurang Q, Cypern, or SysterOBror over the years, which always helped to break the monotony of the day.
This includes but is not limited to Samie, Vishnu, Neel, Sahar, Antonios, Germán, Boules, Baptiste, Håkan, Pol, Marie, and Sebastian.
I would additionally like to extend a special thank you to Sebastian Schiessl, for ``taking me under his wing'' when I first arrived at \gls{ISE} and acting as a sort of mentor during my first moths.
Another big thank you goes out to everyone (current and former) at Satya's group at \gls{CMU}, for always welcoming me with open arms and treating me like one of them whenever I was visiting.
This includes but is not limited to Junjue, Roger, Shilpa, Tom, Jan, Edmond, Jim, Babu, and Zhuo.

I would like to extend my sincere thanks to Prof. Sandra Céspedes, of Universidad de Chile and Concordia University, for her guidance and urging to embark on this journey towards a Ph.D., and for always being willing to offer help and advice.
Likewise, I'd like to thank Prof. Javier Bustos, of Universidad de Chile and NIC Chile Research Labs, for giving me an opportunity many years ago to get into distributed systems research at NIC Labs.
Also, I'd like to extend my gratitude to Prof. José Miguel Piquer, of Universidad de Chile, for taking me as his ``long-distance'' teaching assistant and allowing me to give back to my \emph{alma mater} from across the ocean.

\medskip

This dissertation would not have been possible without the help and support of my family.
First and foremost, I would like to thank the love of my life and wife~\footnote{\emph{future-wife} at the time of writing, but \emph{wife} at the time of the defense of this thesis}, Andresa.
I could not have done this without her companionship, love, and unconditional support.\todo{TODO}

I am also deeply grateful to my parents, Gabriel and Valeria, for always encouraging my curiosity and my drive to learn and discover.
They have been my rocks, my pillars, since the day of my birth, and I am proud to be their son.
A big thank you goes out to my sister, Paola, as well.
Although we sometimes get on each other's nerves, she has always been there for me and supported me, and I could not have asked for a better sister.